\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2020-12-23 manuscript.Rnw
Tweaks to intro

\pkg{hesim} is an \proglang{R} package that advances the functionality and performance of the existing software. Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R}, by inputting parameters obtained from external sources. So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized. After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.

The remainder of this article is organized as follows. Section~\ref{sec:model-taxonomy} describes the economic models supported by \pkg{hesim}. An overview of the coding framework is provided in Section~\ref{sec:framework}. An illustrative example using a three state model of disease progression in oncology is provided in Section~\ref{sec:illustrative-example}. Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient level and aggregate level data. The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples based on the output from the prior section. Section~\ref{sec:discussion} makes comparisons to other software and discusses possible extensions. Finally, Section~\ref{sec:conclusion} concludes.

Analyses were performed using \pkg{hesim} version 0.5.0.

>>>>

\pkg{hesim} is an \proglang{R} package that advances the functionality and performance of the existing software. Multiple model types are supported including cohort discrete time state transition models (cDTSTMs), N-state PSMs, and individual-level continuous time state transition models (iCTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources. So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized. After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.

The remainder of this paper is organized as follows. Section~\ref{sec:model-taxonomy} provides a mathematical description the economic models supported by \pkg{hesim}. An overview of the coding framework is provided in Section~\ref{sec:framework}. An illustrative example using a three state model of disease progression in oncology is provided in Section~\ref{sec:illustrative-example}. Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient-level and aggregate-level data. The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples based on the output from the prior section. Section~\ref{sec:discussion} discusses additional features not covered in this paper, makes comparisons to other software, and explores possible extensions. Finally, Section~\ref{sec:conclusion} concludes.

Analyses were performed using \pkg{hesim} version \code{0.5.0}.
